---
title: "Individual Assignment - BAA5063 Business Statistics Using R"
author: "ARITRA PAUL - Student ID: 25049008"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document: default
  word_document: default
---
# 01. Load Libraries.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)      # Includes dplyr, ggplot2, readr, forcats, etc.
library(janitor)        # For clean_names()
library(caret)          # For createDataPartition() and ML utilities
library(randomForest)   # For random forest model
library(reshape2)       # For melt() function (used in heatmap)
library(GGally)         # For ggpairs
library(factoextra)     # For clustering visualization (fviz_nbclust)
library(knitr)          # For neat table output (kable)
library(scales)         # For comma labels in plots
set.seed(123)
```

# 02. Read data
```{r read-data}
# Read the uploaded CSV
df <- read_csv("C:/Users/aritr/OneDrive/Sunway Studies/BAA5063-Business Statistics Using R/Individual Assignment/Final Submission/cardekho_dataset.csv")
```
# 03. Basic checks
```{r}
glimpse(df)
summary(df)
dim(df)
```
# 04. Missing Value Summary
```{r}
missing_report <- colSums(is.na(df))
missing_report
```

# 05. Clean Data
```{r}
df <- df %>%
  rename(
    price = selling_price,
    fuel = fuel_type,
    transmission = transmission_type
  ) %>%
  mutate(
    age = as.integer(vehicle_age),
    price_per_km = price / (km_driven + 1)
  )

# Trim unnecessary column names
df <- df %>% select(-`...1`)

glimpse(df)
```

# 06. Filter The Data
```{r}
df <- df %>%
filter(!is.na(price), !is.na(km_driven)) %>%
  filter(price > 50000 & price < 5000000)
```

# 07. Feature Engineering
```{r}
df <- df %>%
  mutate(
    brand = fct_lump_min(as.factor(brand), min = 20),
    model = fct_lump_n(as.factor(model), n = 30),
    fuel = as.factor(fuel),
    transmission = as.factor(transmission),
    seller_type = as.factor(seller_type),
    age_group = case_when(
      age <= 3 ~ "0-3",
      age <= 7 ~ "4-7",
      TRUE ~ "8+"
    ),
    km_per_year = km_driven / pmax(age, 0.5)
  )
glimpse(df)
```

# 08. Create a Tidy Numeric-Only df for Correlation / Plotting
```{r}
num_df <- df %>% select(where(is.numeric)) %>% select_if(~sd(., na.rm=TRUE) != 0)
glimpse(num_df)
```

# 09. Exploratory Data Analysis

## 09.1. Histogram of Distibution of Listing Prices

```{r}
ggplot(df, aes(x = price)) +
  geom_histogram(bins = 80 ,fill="lightblue" , col = "black") +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Distribution of Listing Prices", x = "Selling Price (INR)", y = "Count")+
  theme_minimal()
```

## 09.2. Top 10 Brands by Average Price

```{r}
top_brands <- df %>%
  group_by(brand) %>%
  summarise(n = n(), avg_price = mean(price, na.rm = TRUE)) %>%
  arrange(desc(n)) %>%
  slice(1:10)

ggplot(top_brands, aes(x = reorder(brand, avg_price), y = avg_price)) +
  geom_col(fill="darkolivegreen2") +
  coord_flip() + 
  labs(title = "Top 10 Brands by Average Price", x = "Brand", y = "Average Price")+
  theme_minimal()
```

## 09.3. Box Plot (Selling Price by Fuel Type)

```{r fig.width=10, fig.height=5}
ggplot(df, aes(x = fuel, y = price)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 100)+
  scale_y_continuous(labels =scales::comma) + labs(title = "Price of Cars as Per Fuel Type", x = "Fuel Type", y="Selling Price (INR)")+
  theme_minimal()
```

## 09.4. Price Vs Age

```{r}
ggplot(df, aes(x = age, y = price)) +
  geom_point(alpha = 0.3, col = "hotpink4") +
  geom_smooth(method = "loess") +
  scale_y_continuous(labels =scales::comma)+
  labs(title = "Selling Price of Car With Respect to Age", x = "Age (years)", y = "Selling Price (INR)")+
  theme_minimal()
```

## 09.5. Price Vs Age (Based on Transmission)

```{r}
ggplot(df, aes(x = vehicle_age, y = price, color = transmission)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") + 
  scale_y_continuous(labels =scales::comma) + labs(title = "Price of Cars as According to Age of Vehicle WRT Transmission", x = "Age of Vehicle", y= "Selling Price (INR)",color="Transmission")+
  theme_minimal()
```

## 09.6. Correlation Matrix & Heatmap

```{r}
# Correlation matrix
cor_matrix <- cor(num_df, use = "complete.obs")
cor_matrix

# Reshape for ggplot
cor_m <- melt(cor_matrix)


# Enhanced heatmap with values and size control

heatMap <- ggplot(cor_m, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  
  # Add correlation values as text
  geom_text(aes(label = round(value, 2)), color = "black", size = 3.5) +
  
  # Custom color palette
  scale_fill_gradientn(colors = c("violetred1", "lavender", "turquoise"),
                       limits = c(-1, 1),
                       name = "Correlation") +
  
  # Improve readability
  theme_minimal(base_size = 13.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  labs(title = "Correlation Heatmap", x = "", y = "")

heatMap
```

```{r fig.width=14, fig.height=10}
ggpairs_df <- df %>% 
  select(price, age, km_driven, km_per_year,transmission) %>%
  drop_na()

# Create the ggpairs plot
ggpairs(
  ggpairs_df,
  columns = 1:4,
  aes(color = transmission, alpha = 0.6),
  upper = list(continuous = wrap("cor", size = 5)),   # show correlation in upper
  lower = list(continuous = wrap("points", alpha = 0.4, size = 1)), # scatter below
  diag = list(continuous = wrap("densityDiag"))       # density plots on diagonal
) +
  theme_bw() +
  theme(
    text = element_text(size = 18),
    axis.text.x = element_text(angle = 90, hjust = 1)
  )+ scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma)
```

# 10. Train-Test Split

```{r}
train_index <- createDataPartition(df$price, p = 0.8, list = FALSE)
train <- df[train_index, ]
test  <- df[-train_index, ]

cols_to_fix <- c("fuel", "transmission", "seller_type", "brand", "model")
for (col in cols_to_fix) {
  test[[col]] <- factor(test[[col]], levels = levels(train[[col]]))
}
```

# 11. Linear Regression on Train Set

## 11.1. Model 1
```{r}
model1 <- lm(
  log(price) ~ age,
  data = train)

summary(model1)

```

## 11.2. Model 2

```{r}
model2 <- lm(
  log(price) ~ age + log(km_driven + 1),
  data = train)

summary(model2)

```

## 11.3. Model 3

```{r}
model3 <- lm(
  log(price) ~ age + log(km_driven + 1) + fuel,
  data = train)

summary(model3)

```

## 11.4. Model 4

```{r}
model4 <- lm(
  log(price) ~ age + log(km_driven + 1) + fuel + transmission,
  data = train)

summary(model4)

```

## 11.5. Model 5

```{r}
model5 <- lm(
  log(price) ~ age + log(km_driven + 1) + fuel + transmission + seller_type ,
  data = train)

summary(model5)

```

## 11.6. Model 6

```{r}
model6 <- lm(
  log(price) ~ age + log(km_driven + 1) + fuel + transmission + seller_type + brand ,
  data = train)

summary(model6)

```

## 11.7. Model 7

```{r}
model7 <- lm(
  log(price) ~ age + log(km_driven + 1) + fuel + transmission + seller_type + brand + model,
  data = train)

summary(model7)
```
```{r}
# Compare R² values

c(summary(model1)$r.squared,
  summary(model2)$r.squared,
  summary(model3)$r.squared,
  summary(model4)$r.squared,
  summary(model5)$r.squared,
  summary(model6)$r.squared,
  summary(model7)$r.squared)
```

# 12. Diagnostics

```{r}
plot(model7, col="yellow3", which = c(1))
plot(model7, col="red", which = c(2))
plot(model7, col="skyblue", which = c(3))
plot(model7, col="red", which = c(4))

```

# 13. Prediction on Test Set (Linear Regression)

```{r}
pred_lm <- exp(predict(model7, newdata = test))

lm_rmse <- sqrt(mean((pred_lm - test$price)^2, na.rm = TRUE))
lm_mae  <- mean(abs(pred_lm - test$price), na.rm = TRUE)
lm_r2   <- 1 - sum((pred_lm - test$price)^2, na.rm = TRUE) /
                sum((mean(train$price, na.rm = TRUE) - test$price)^2, na.rm = TRUE)

list(RMSE = lm_rmse, MAE = lm_mae, r2 = lm_r2)
```

# 14. Training, and Prediction on Test Set (Random Forest)

```{r}
train_rf <- train %>% sample_n(min(nrow(train), 15000))

rf <- randomForest(
   price ~ age + km_driven+ fuel + transmission + seller_type + brand + model,
  data = train_rf,
  ntree = 200,
  importance = TRUE
)

rf_pred <- predict(rf, newdata = test)

rf_rmse <- sqrt(mean((rf_pred - test$price)^2, na.rm = TRUE))
rf_mae  <- mean(abs(rf_pred - test$price), na.rm = TRUE)
rf_r2   <- 1 - sum((rf_pred - test$price)^2, na.rm = TRUE) /
                sum((mean(train$price, na.rm = TRUE) - test$price)^2, na.rm = TRUE)

list(rmse = rf_rmse, mae = rf_mae, r2 = rf_r2)

```

# 15. Model Performance Comparison

```{r}
results <- tibble(
  Model = c("Linear Regression", "Random Forest"),
  RMSE = c(lm_rmse, rf_rmse),
  MAE = c(lm_mae, rf_mae),
  R2 = c(lm_r2, rf_r2)
)
knitr::kable(results, caption = "**Model Performance Comparison**")
```

# 16. Variable Importance

```{r}
# Variable Importance - Random Forest
options(scipen = 999)
importance(rf)
varImpPlot(rf, main = "Variable Importance", col = "red", pch = 18)
```


# 17. Clustering

```{r}
clu_df <- df %>% select(price, age, km_per_year) %>% drop_na()
clu_scaled <- scale(clu_df)
fviz_nbclust(clu_scaled, kmeans, method = "wss")
km <- kmeans(clu_scaled, centers = 4, nstart = 25)

# Assigning Cluster Labels
clu_df$cluster <- factor(km$cluster)

# Cluster Counts
table(clu_df$cluster)
```

